---
title: "Stats Project"
output: html_document
---

In this project, we're using sales data from Safeway Eastern's floral department. This data only shows the flowers/bouquets sold at Safeway Eastern supermarkets that were bought from Intergreen USA, a flower importing company which supplies different supermarkets across the US. There are a total of 9 distinct products sold and the sales occurred between April and October 2017. 

We're interested in seeing the relationships between several predictor variables and sales to determine if there's a particular independent variable that drives or hurts flower sales. 
One of these variables is the closing price of the S&P 500 for a given day. We chose this variable because the S&P 500 includes stocks from large companies that sell consumer staples like Johnson and Johnson, Heinz, Colgate, and Coca-Cola. Many of their products are sold in supermarkets, so our reasoning is that the S&P will reflect the movement in supermarkets. Similarly, since the flowers in the data are being sold in supermarkets this could be a good predictor for flower sales. 


The code below imports the flower sales data and combines it with the S&P data.
```{r cars}
library('plyr')
library('bit64')
library('dplyr')
library('reshape2')
library('ggplot2')
library('data.table')

setwd("C:/Users/Holex/Desktop/Stats_Project")

#Sidenote: I saved excel file as a txt before reading it as a table
flower_data <- read.table("flower_data.txt", header = TRUE)
flower_data_apr <- fread("april_june.csv", header = TRUE)


#line below ensures column is date and ordered
flower_data$Txn_Date <- as.Date(flower_data$Txn_Date, format = "%m/%d/%Y")
flower_data_apr$Day <- as.Date(flower_data_apr$Day, format = "%m/%d/%Y")


#Since there are multiple sales on a given day, I want to aggregate the sales based on the date
date_net <- aggregate(Net ~ Txn_Date, flower_data,sum)
colnames(flower_data_apr)[6] <- "Net"
date2_net <- aggregate(Net ~ Day, flower_data_apr,sum)

#renaming columns to make joining easier later
#colnames(date2_net)[2] <- "Net"
colnames(date2_net)[1] <- "Date"

#read in financial data
stock_info <- fread("stockdata.csv")

stock_info2 <- fread("april_june_finance.csv")
stock_info2$Date <- as.Date(stock_info2$Date, format = "%m/%d/%Y")
#change column name so i can do a join
colnames(date_net)[1] <- "Date"
combodata <- join(date_net,stock_info, type="right")
combodata2 <- join(date2_net, stock_info2, type = "right")
combodata2$Date <- as.character(combodata2$Date)
#combo3 <- merge(x = date2_net, y = stock_info2, by = "Date", all= TRUE) 

#bind 2 data frames (data from april-june and sep-oct)
results <- rbind(combodata2, combodata)

```
Next, we looked at plots of flower sales against the S&P open, close, high, adjusted close, low, and volume individually to see which plots looked linear and could make it to the next stage of the data analysis process. 

```{r}
plot(results$Open, results$Net)
plot(results$Close, results$Net)
plot(results$High, results$Net)
plot(results$`Adj Close`, results$Net)
plot(results$Low, results$Net)
plot(results$Volume, results$Net)
```



```{r}
l.model <- lm(results$Net ~ results$`Adj Close`)
summary(l.model)
#plot against residuals
plot(l.model$fitted.values, l.model$residuals)
qqnorm(rstandard(l.model))
qqline(rstandard(l.model))
plot(l.model)
#pollen count for day try other predictor variables
```

#Testing code here

```{r}
#date_net3 <- aggregate(Net ~ Txn_Date, flower_data,sum)
#date_net3$Txn_Date <- factor(date_net3$Txn_Date, levels =  #date_net3$Txn_Date[order(date_net3$Net)])
#str(date_net3)

#date_net3$Txn_Date <- format(format.Date(date_net3))

#There are also multiple stores, it would be interesting to see how different stores sell based on the day
#link to safeway eastern stores
#https://www.google.com/maps/d/viewer?mid=1SU7aT0QrKqPL-5mGKjwvflXmAz4&hl=en_US&ll=39.02715642805036%2C-76.56616499999996&z=7


store_date_net2 <- aggregate(Net ~ Txn_Date + Store , flower_data, sum)
store_date_net2 <- store_date_net[order(as.Date(store_date_net2$Txn_Date, "%m/%d/%Y")),]


```

```{r}

library('rnoaa')
#https://cran.r-project.org/web/packages/rnoaa/rnoaa.pdf

isd_stations_search(lat=38.901335, lon=-76.980425)

#code to get station name for every longitude and latitude for Safeway Eastern stores


store_locations <- read.table("lat_lng_store.txt", header = TRUE)


coops_search(station_name = 997314, begin_date = 20140928, end_date = 20140929, datum = "stnd", product = "predictions")

ghcnd_search("99999", date_min = "1920-01-01", date_max = "1925-01-01") 

#create a loop that stores dont neeed loop can pass a vector ;)

#meteo_tidy_ghcnd(stationid = "ASN00003003", date_min = "1950-01-01")

station_data <- ghcnd_stations()

meteo_distance(station_data, lat, long, units = "deg", radius = NULL, limit = NULL)
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}

meteo_distance(station_data, 38.901335, -76.98042, units = "deg", radius = NULL, limit = 5)
```

Stock market data
```{r}
library('quantmod')
```

